---
title: "WGCNA_TG_DM_rat_kidney"
author:
  - name: "Steven J. Kunnen"
    url: "https://orcid.org/0000-0001-9549-1719"
    affiliation: "Leiden University"
    affiliation_url: "https://ror.org/027bh9e22"
date: "`r Sys.Date()`"
output: html_document
---


# Setup
```{r setup}
# Setup of WGCNA input parameters and files
username = Sys.getenv("USERNAME")
softpower = 8
min_module_size = 5
deepsplit = 4
cutheight = 0.2 # Corresponds to a correlation threshold of 0.8 (1 - 0.8 = 0.2)
verbose = 3
network_type = "unsigned"
method = "average"
expression_file = "RG230-2.kidney_13582geneExpression_DM_TGexprs_brainarray.csv"
```


# Load libraries
```{r, message=FALSE}
if(!require("BiocManager", quietly = TRUE)) install.packages("BiocManager") 
library(BiocManager)

BiocManager::install(c("org.Hs.eg.db", "org.Mm.eg.db", "GO.db", "impute", "minet", "preprocessCore"))

if(!require("pacman", quietly = TRUE)) install.packages("pacman") 
library(pacman)

p_load(tidyverse, data.table, parallel, patchwork, readxl, WGCNA, dynamicTreeCut, fastcluster, flashClust, RColorBrewer, devtools, plotly)

if(WGCNA::WGCNAnThreads()==1) {
  WGCNA::enableWGCNAThreads(nThreads = floor(parallel::detectCores() * 0.8))
}

inputDir <- file.path(getwd(), "DATA")

outputDir <- file.path(getwd(), "OUTPUT")

# create data_id output folder
if(!dir.exists(outputDir)){
  dir.create(path = outputDir)
}

session_info()
```


# Load data: gene expression log2FC matrix (genes in rows, samples/conditions in columns)
```{r}
# Step 1: Load your gene expression data
# Assume the data is in a CSV file (rows: genes, columns: samples)
input <- read.csv(file.path(inputDir, expression_file), header = T)

experiment_info <- read_xlsx(file.path(inputDir, "kidney_experiments.xlsx"))

expression_data <- input |>
  mutate(entrez_id = paste0("id_", geneid)) |>
  dplyr::select(entrez_id, contains("EXP_")) |>
  column_to_rownames("entrez_id") |>
  t() |>
  as.data.frame()

# Step 2: Data Cleaning
# Remove genes and samples with too many missing values
goodSamplesGenes <- goodSamplesGenes(expression_data, verbose = verbose)

expression_data <- expression_data[goodSamplesGenes$goodSamples, goodSamplesGenes$goodGenes]

# Step 3: Z-scale the data (center and scale)
expression_data_zscaled <- data.frame(scale(x = expression_data, center = T, scale = T))
expression_data_zscaled_check <- data.frame(apply(expression_data, 2, function(x) (x-mean(x))/sd(x)))
```


# WGCNA TG + DM combined

## 1a. Soft-threshold Selection: This step helps select an appropriate power value to ensure the network follows a scale-free topology.
```{r}
powers <- c(1:20)

# Calculate scale-free topology fit using soft power range
soft_data <- WGCNA::pickSoftThreshold(data = expression_data,
                                       powerVector = powers,
                                       verbose = verbose,
                                       blockSize = 2000,
                                       networkType = network_type,
                                       nBreaks = 10) 

soft_data_zscaled <- WGCNA::pickSoftThreshold(data = expression_data_zscaled,
                                               powerVector = powers,
                                               verbose = verbose,
                                               blockSize = 2000,
                                               networkType = network_type,
                                               nBreaks = 10) 

soft_data$fitIndices <- soft_data$fitIndices |>
    dplyr::mutate(r = -sign(slope) * SFT.R.sq)

soft_data_zscaled$fitIndices <- soft_data$fitIndices |>
    dplyr::mutate(r = -sign(slope) * SFT.R.sq)

# Plot the results to decide the best soft threshold
par(mfrow = c(1, 2))
plot(soft_data$fitIndices[, 1], -sign(soft_data$fitIndices[, 3]) * soft_data$fitIndices[, 2],
     xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n", main = "Scale independence")
text(soft_data$fitIndices[, 1], -sign(soft_data$fitIndices[, 3]) * soft_data$fitIndices[, 2],
     labels = powers, cex = 0.9, col = "red")
abline(h = 0.90, col = "red") # 0.90 scale-free topology criterion

# Mean connectivity as a function of soft-threshold power
plot(soft_data$fitIndices[, 1], soft_data$fitIndices[, 5],
     xlab = "Soft Threshold (power)", ylab = "Mean Connectivity", type = "n",
     main = "Mean connectivity")
text(soft_data$fitIndices[, 1], soft_data$fitIndices[, 5], labels = powers, cex = 0.9, col = "red")

# plot 1: scale independence scatter plot
p1 <- ggplot(data = soft_data$fitIndices, mapping = aes(x = Power, y = r, label = Power)) +
  geom_text(color = "red") +
  geom_hline(yintercept = 0.9,color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.1), limits = c(0, 1)) +
  labs(title = paste0("Scale independance"), x = "Softpower", y = "Scale free topology model fit") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

p3 <- ggplot(data = soft_data_zscaled$fitIndices, mapping = aes(x = Power, y = r, label = Power)) +
  geom_text(color = "red") +
  geom_hline(yintercept = 0.9,color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.1), limits = c(0, 1)) +
  labs(title = paste0("Scale independance"), x = "Softpower", y = "Scale free topology model fit") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

# plot 2: mean connectivity
p2 <- ggplot(data = soft_data$fitIndices , mapping = aes(x = Power, y = mean.k., label  = Power)) +
  geom_text(color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(transform = "log10", breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) +
  labs(title = paste0("Mean connectivity"), x = "Softpower", y = "Mean connectivity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

p4 <- ggplot(data = soft_data_zscaled$fitIndices , mapping = aes(x = Power, y = mean.k., label  = Power)) +
  geom_text(color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(transform = "log10", breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) +
  labs(title = paste0("Mean connectivity"), x = "Softpower", y = "Mean connectivity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))


(p1|p2)/(p3|p4)

ggsave(paste0(outputDir,"/Softpower_plot_TG_DM_",min(powers),"_till_",max(powers),".pdf"), width = 10, height = 10)

# Difference between scale-free topology fit based on normal log2FC or zscaled log2FC
setdiff(soft_data$fitIndices,soft_data_zscaled$fitIndices)

# Select of adjust the softpower used to run the WGCNA model.
softpower = 8 # Modify based on the scale free topology plot results
```


## 2a. Construct the Network and hierarchical clustering
```{r}
# Adjacency matrix using the soft-thresholding power
message("Softpower: ", softpower)

message("Adjacency matrix")
adjacency_matrix = WGCNA::adjacency(datExpr = expression_data,
                                    type = network_type,
                                    power = softpower)


# Topological Overlap Matrix (TOM)
message("TOM")
tom <- WGCNA::TOMsimilarity(adjMat = adjacency_matrix,
                            TOMType = network_type,
                            verbose = verbose)

message("disTOM")
diss_tom <- 1 - tom

# Cluster genes using hierarchical clustering
message("flashClust")
geneTree <- fastcluster::hclust(as.dist(diss_tom), method = method)
dendro <- flashClust::flashClust(as.dist(diss_tom), method = method)

plot(geneTree, main = "Gene Clustering on TOM-based Dissimilarity", sub = "", xlab = "", cex = 0.6)
plot(dendro, main = "Gene Clustering on TOM-based Dissimilarity", sub = "", xlab = "", cex = 0.6)

identical(round(geneTree$height,3),round(dendro$height,3))
```


## 3a. Module Identification using Dynamic Tree Cut
```{r}
# Step 1: Module Identification using Dynamic Tree Cut
# Use dynamic tree cut to identify modules
message("cutreeDynamic")

dynamicMods <- cutreeDynamic(dendro = geneTree, 
                             distM = diss_tom,
                             deepSplit = deepsplit, 
                             pamRespectsDendro = FALSE,
                             minClusterSize = min_module_size,
                             verbose = verbose)

module_table_premerge <- as.data.frame(table(dynamicMods))

# Convert numeric labels into colors for visualization
dynamicColors <- labels2colors(dynamicMods)
table(dynamicColors)

# Plot dendrogram with module colors
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)


# Step 2: Calculate Module Eigengenes
MEs <- moduleEigengenes(expression_data, colors = dynamicMods)$eigengenes

# Calculate the dissimilarity of module eigengenes
ME_diss <- 1 - cor(MEs)

# Hierarchical clustering of module eigengenes
ME_tree <- hclust(as.dist(ME_diss), method = "average")
plot(ME_tree, main = "Clustering of Module Eigengenes", xlab = "", sub = "")


# Step 3: Merge Similar Modules (module eigengenes with correlation > 0.8)
message("mergeCloseModules")

mergedModules <- mergeCloseModules(exprData = expression_data, 
                                   colors = dynamicMods, 
                                   cutHeight = cutheight,
                                   verbose = verbose,
                                   iterate = FALSE)

# Get the new merged module colors
mergedColors <- mergedModules$colors

module_table_postmerge <- as.data.frame(table(mergedColors))

# Get the new merged module eigengenes
mergedMEs <- mergedModules$newMEs

# Plot dendrogram with merged module colors
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
                    c("Original Modules", "Merged Modules"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)

eg_score_mergedMEs_summary = mergedMEs |> summary() |> as.data.frame() |> select(-Var1) |> 
  separate(col = Freq, into = c("Stat", "Value"), sep = ":") |> 
  pivot_wider(names_from = Stat, values_from = Value)


# Step 4: Get module definition
module_definition <- data.frame(module_premerge = paste0("module_",dynamicMods),
                                module_postmerge = paste0("module_",mergedColors),
                                entrez_id = colnames(expression_data),
                                softpower = softpower,
                                deepsplit = deepsplit,
                                minclustersize = min_module_size,
                                cutheight = cutheight) |>
  dplyr::left_join(y = input |>
                     mutate(entrez_id = paste0("id_", geneid)) |> 
                     dplyr::select(gene_symbol, entrez_id), by = "entrez_id")

module_table_merging <- module_definition |> 
  select(module_premerge, module_postmerge) |>
  mutate(module_nr = as.integer(str_remove_all(module_postmerge, "module_"))) |> 
  table() |> as.data.frame() |> 
  filter(Freq > 0) 


# Step 5: Z-score module eigengene scores premerge
MEs_old_scaled <- data.frame(apply(mergedModules$oldMEs, 2, function(x) (x-mean(x))/sd(x))) 

Cor_MEs_old <- mergedModules$oldMEs |>
  # column_to_rownames("sample_id") |>
  cor() |>
  as.data.frame() |>
  rownames_to_column("module") |>
  mutate(module = paste0("module_", module)) |>
  pivot_longer(cols = where(is.numeric), names_to = "module_cor", values_to = "PearsonR") 

Cor_MEs_old_scaled <- MEs_old_scaled |>
  # column_to_rownames("sample_id") |>
  cor() |>
  as.data.frame() |>
  rownames_to_column("module") |>
  mutate(module = paste0("module_", module)) |>
  pivot_longer(cols = where(is.numeric), names_to = "module_cor", values_to = "PearsonR") 

Cor_MEs_old_merge <- Cor_MEs_old |> filter(PearsonR > 0.80 & PearsonR < 0.99)

Cor_MEs_old_scaled_merge <- Cor_MEs_old_scaled |> filter(PearsonR > 0.80 & PearsonR < 0.99)


if(!mergedModules$allOK) {
  return()
}

# Step 6: Scale module eigengene scores
message("Module eigengene score premerge")
colnames(mergedModules$oldMEs) <- gsub("ME", "", colnames(mergedModules$oldMEs))

eg_score_premerge <- mergedModules$oldMEs |>
  tibble::rownames_to_column(var = "sample_id")

eg_score_premerge_scaled <- data.frame(apply(mergedModules$oldMEs, 2, function(x) x/sd(x))) |>
  tibble::rownames_to_column(var = "sample_id")

message("Module eigengene score postmerge")
colnames(mergedModules$newMEs) <- gsub("ME", "", colnames(mergedModules$newMEs))

eg_score_postmerge <- mergedModules$newMEs |>
  tibble::rownames_to_column(var = "sample_id")

eg_score_postmerge_scaled <- data.frame(apply(mergedModules$newMEs, 2, function(x) x/sd(x))) |>
  tibble::rownames_to_column(var = "sample_id")

eg_score_postmerge_summary = eg_score_postmerge |> select(-sample_id) |> summary() |> as.data.frame() |> select(-Var1) |> 
  separate(col = Freq, into = c("Stat", "Value"), sep = ":") |> 
  pivot_wider(names_from = Stat, values_from = Value)

eg_score_postmerge_scaled_summary = eg_score_postmerge_scaled |> select(-sample_id)|> summary() |> as.data.frame() |> select(-Var1) |> 
  separate(col = Freq, into = c("Stat", "Value"), sep = ":") |> 
  pivot_wider(names_from = Stat, values_from = Value)


Cor_MEs_new <- eg_score_postmerge |>
  column_to_rownames("sample_id") |>
  cor() |>
  as.data.frame() |>
  rownames_to_column("module") |>
  mutate(module = paste0("module_", module)) |>
  pivot_longer(cols = where(is.numeric), names_to = "module_cor", values_to = "PearsonR") 
  
Cor_MEs_new_merge <- Cor_MEs_new |> filter(PearsonR > 0.80 & PearsonR < 0.99)
```


# WGCNA TG data

## 1b. TG: Soft-threshold Selection: This step helps select an appropriate power value to ensure the network follows a scale-free topology.
```{r}
# Subset expression_data
experiment_TG = experiment_info |>
  filter(Source == "TGGates", TISSUE_NAME == "KIDNEY") |>
  mutate(EXP_NAME = paste0("EXP_", EXPERIMENT))
  
expression_data_TG = expression_data |> 
  rownames_to_column("experiment") |>
  filter(experiment %in% experiment_TG$EXP_NAME) |>
  column_to_rownames("experiment") 

expression_data_zscaled_TG <- data.frame(scale(x = expression_data_TG, center = T, scale = T))

powers <- c(1:20)

# Calculate scale-free topology fit using soft power range
soft_data_TG <- WGCNA::pickSoftThreshold(data = expression_data_TG,
                                       powerVector = powers,
                                       verbose = verbose,
                                       blockSize = 2000,
                                       networkType = network_type,
                                       nBreaks = 10) 

soft_data_zscaled_TG <- WGCNA::pickSoftThreshold(data = expression_data_zscaled_TG,
                                               powerVector = powers,
                                               verbose = verbose,
                                               blockSize = 2000,
                                               networkType = network_type,
                                               nBreaks = 10) 

soft_data_TG$fitIndices <- soft_data_TG$fitIndices |>
    dplyr::mutate(r = -sign(slope) * SFT.R.sq)

soft_data_zscaled_TG$fitIndices <- soft_data_TG$fitIndices |>
    dplyr::mutate(r = -sign(slope) * SFT.R.sq)

# Plot the results to decide the best soft threshold
par(mfrow = c(1, 2))
plot(soft_data_TG$fitIndices[, 1], -sign(soft_data_TG$fitIndices[, 3]) * soft_data_TG$fitIndices[, 2],
     xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n", main = "Scale independence")
text(soft_data_TG$fitIndices[, 1], -sign(soft_data_TG$fitIndices[, 3]) * soft_data_TG$fitIndices[, 2],
     labels = powers, cex = 0.9, col = "red")
abline(h = 0.90, col = "red") # 0.90 scale-free topology criterion

# Mean connectivity as a function of soft-threshold power
plot(soft_data_TG$fitIndices[, 1], soft_data_TG$fitIndices[, 5],
     xlab = "Soft Threshold (power)", ylab = "Mean Connectivity", type = "n",
     main = "Mean connectivity")
text(soft_data_TG$fitIndices[, 1], soft_data_TG$fitIndices[, 5], labels = powers, cex = 0.9, col = "red")

# plot 1: scale independence scatter plot
p1 <- ggplot(data = soft_data_TG$fitIndices, mapping = aes(x = Power, y = r, label = Power)) +
  geom_text(color = "red") +
  geom_hline(yintercept = 0.9,color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.1), limits = c(0, 1)) +
  labs(title = paste0("Scale independance"), x = "Softpower", y = "Scale free topology model fit") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

p3 <- ggplot(data = soft_data_zscaled_TG$fitIndices, mapping = aes(x = Power, y = r, label = Power)) +
  geom_text(color = "red") +
  geom_hline(yintercept = 0.9,color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.1), limits = c(0, 1)) +
  labs(title = paste0("Scale independance"), x = "Softpower", y = "Scale free topology model fit") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

# plot 2: mean connectivity
p2 <- ggplot(data = soft_data_TG$fitIndices , mapping = aes(x = Power, y = mean.k., label  = Power)) +
  geom_text(color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(transform = "log10", breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) +
  labs(title = paste0("Mean connectivity"), x = "Softpower", y = "Mean connectivity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

p4 <- ggplot(data = soft_data_zscaled_TG$fitIndices , mapping = aes(x = Power, y = mean.k., label  = Power)) +
  geom_text(color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(transform = "log10", breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) +
  labs(title = paste0("Mean connectivity"), x = "Softpower", y = "Mean connectivity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))


(p1|p2)/(p3|p4)

ggsave(paste0(outputDir,"/Softpower_plot_TG_",min(powers),"_till_",max(powers),".pdf"), width = 10, height = 10)

# Difference between scale-free topology fit based on normal log2FC or zscaled log2FC
setdiff(soft_data_TG$fitIndices,soft_data_zscaled_TG$fitIndices)

# Select of adjust the softpower used to run the WGCNA model.
softpower = 8 # Modify based on the scale free topology plot results
```


## 2b. TG: Construct the Network and hierarchical clustering
```{r}
# Adjacency matrix using the soft-thresholding power
message("Softpower: ", softpower)

message("Adjacency matrix")
adjacency_matrix_TG = WGCNA::adjacency(datExpr = expression_data_TG,
                                    type = network_type,
                                    power = softpower)


# Topological Overlap Matrix (TOM)
message("TOM")
tom_TG <- WGCNA::TOMsimilarity(adjMat = adjacency_matrix_TG,
                            TOMType = network_type,
                            verbose = verbose)

message("disTOM")
diss_tom_TG <- 1 - tom_TG

# Cluster genes using hierarchical clustering
message("flashClust")
geneTree_TG <- fastcluster::hclust(as.dist(diss_tom_TG), method = method)
dendro_TG <- flashClust::flashClust(as.dist(diss_tom_TG), method = method)

plot(geneTree_TG, main = "Gene Clustering on TOM-based Dissimilarity", sub = "", xlab = "", cex = 0.6)
plot(dendro_TG, main = "Gene Clustering on TOM-based Dissimilarity", sub = "", xlab = "", cex = 0.6)

identical(round(geneTree_TG$height,3),round(dendro_TG$height,3))
```


## 3b. TG: Module Identification using Dynamic Tree Cut
```{r}
# Step 1: Module Identification using Dynamic Tree Cut
# Use dynamic tree cut to identify modules
message("cutreeDynamic")

dynamicMods_TG <- cutreeDynamic(dendro = geneTree_TG, 
                             distM = diss_tom_TG,
                             deepSplit = deepsplit, 
                             pamRespectsDendro = FALSE,
                             minClusterSize = min_module_size,
                             verbose = verbose)

module_table_premerge_TG <- as.data.frame(table(dynamicMods_TG))

# Convert numeric labels into colors for visualization
dynamicColors_TG <- labels2colors(dynamicMods_TG)
table(dynamicColors_TG)

# Plot dendrogram with module colors
plotDendroAndColors(geneTree, dynamicColors_TG, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)


# Step 2: Calculate Module Eigengenes
MEs_TG <- moduleEigengenes(expression_data_TG, colors = dynamicMods_TG)$eigengenes

# Calculate the dissimilarity of module eigengenes
ME_diss_TG <- 1 - cor(MEs_TG)

# Hierarchical clustering of module eigengenes
ME_tree_TG <- hclust(as.dist(ME_diss_TG), method = "average")
plot(ME_tree_TG, main = "Clustering of Module Eigengenes", xlab = "", sub = "")


# Step 3: Merge Similar Modules (module eigengenes with correlation > 0.8)
message("mergeCloseModules")

mergedModules_TG <- mergeCloseModules(exprData = expression_data_TG, 
                                   colors = dynamicMods_TG, 
                                   cutHeight = cutheight,
                                   verbose = verbose,
                                   iterate = FALSE)

# Get the new merged module colors
mergedColors_TG <- mergedModules_TG$colors

module_table_postmerge_TG <- as.data.frame(table(mergedColors_TG))

# Get the new merged module eigengenes
mergedMEs_TG <- mergedModules_TG$newMEs

# Plot dendrogram with merged module colors
plotDendroAndColors(geneTree, cbind(dynamicColors_TG, mergedColors_TG),
                    c("Original Modules", "Merged Modules"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)

eg_score_mergedMEs_summary = mergedMEs_TG |> summary() |> as.data.frame() |> select(-Var1) |> 
  separate(col = Freq, into = c("Stat", "Value"), sep = ":") |> 
  pivot_wider(names_from = Stat, values_from = Value)


# Step 4: Get module definition
module_definition_TG <- data.frame(module_premerge = paste0("module_",dynamicMods_TG),
                                module_postmerge = paste0("module_",mergedColors_TG),
                                entrez_id = colnames(expression_data_TG),
                                softpower = softpower,
                                deepsplit = deepsplit,
                                minclustersize = min_module_size,
                                cutheight = cutheight) |>
  dplyr::left_join(y = input |>
                     mutate(entrez_id = paste0("id_", geneid)) |> 
                     dplyr::select(gene_symbol, entrez_id), by = "entrez_id")

module_table_merging_TG <- module_definition_TG |> 
  select(module_premerge, module_postmerge) |>
  mutate(module_nr = as.integer(str_remove_all(module_postmerge, "module_"))) |> 
  table() |> as.data.frame() |> 
  filter(Freq > 0) 
```


# WGCNA DM data

## 1c. DM: Soft-threshold Selection: This step helps select an appropriate power value to ensure the network follows a scale-free topology.
```{r}
# Subset expression_data
experiment_DM = experiment_info |>
  filter(Source == "DM", TISSUE_NAME == "KIDNEY") |>
  mutate(EXP_NAME = paste0("EXP_", EXPERIMENT))
  
expression_data_DM = expression_data |> 
  rownames_to_column("experiment") |>
  filter(experiment %in% experiment_DM$EXP_NAME) |>
  column_to_rownames("experiment") 

expression_data_zscaled_DM <- data.frame(scale(x = expression_data_DM, center = T, scale = T))

powers <- c(1:20)

# Calculate scale-free topology fit using soft power range
soft_data_DM <- WGCNA::pickSoftThreshold(data = expression_data_DM,
                                       powerVector = powers,
                                       verbose = verbose,
                                       blockSize = 2000,
                                       networkType = network_type,
                                       nBreaks = 10) 

soft_data_zscaled_DM <- WGCNA::pickSoftThreshold(data = expression_data_zscaled_DM,
                                               powerVector = powers,
                                               verbose = verbose,
                                               blockSize = 2000,
                                               networkType = network_type,
                                               nBreaks = 10) 

soft_data_DM$fitIndices <- soft_data_DM$fitIndices |>
    dplyr::mutate(r = -sign(slope) * SFT.R.sq)

soft_data_zscaled_DM$fitIndices <- soft_data_DM$fitIndices |>
    dplyr::mutate(r = -sign(slope) * SFT.R.sq)

# Plot the results to decide the best soft threshold
par(mfrow = c(1, 2))
plot(soft_data_DM$fitIndices[, 1], -sign(soft_data_DM$fitIndices[, 3]) * soft_data_DM$fitIndices[, 2],
     xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n", main = "Scale independence")
text(soft_data_DM$fitIndices[, 1], -sign(soft_data_DM$fitIndices[, 3]) * soft_data_DM$fitIndices[, 2],
     labels = powers, cex = 0.9, col = "red")
abline(h = 0.90, col = "red") # 0.90 scale-free topology criterion

# Mean connectivity as a function of soft-threshold power
plot(soft_data_DM$fitIndices[, 1], soft_data_DM$fitIndices[, 5],
     xlab = "Soft Threshold (power)", ylab = "Mean Connectivity", type = "n",
     main = "Mean connectivity")
text(soft_data_DM$fitIndices[, 1], soft_data_DM$fitIndices[, 5], labels = powers, cex = 0.9, col = "red")

# plot 1: scale independence scatter plot
p1 <- ggplot(data = soft_data_DM$fitIndices, mapping = aes(x = Power, y = r, label = Power)) +
  geom_text(color = "red") +
  geom_hline(yintercept = 0.9,color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.1), limits = c(0, 1)) +
  labs(title = paste0("Scale independance"), x = "Softpower", y = "Scale free topology model fit") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

p3 <- ggplot(data = soft_data_zscaled_DM$fitIndices, mapping = aes(x = Power, y = r, label = Power)) +
  geom_text(color = "red") +
  geom_hline(yintercept = 0.9,color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(breaks = seq(from = 0, to = 1, by = 0.1), limits = c(0, 1)) +
  labs(title = paste0("Scale independance"), x = "Softpower", y = "Scale free topology model fit") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

# plot 2: mean connectivity
p2 <- ggplot(data = soft_data_DM$fitIndices , mapping = aes(x = Power, y = mean.k., label  = Power)) +
  geom_text(color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(transform = "log10", breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) +
  labs(title = paste0("Mean connectivity"), x = "Softpower", y = "Mean connectivity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

p4 <- ggplot(data = soft_data_zscaled_DM$fitIndices , mapping = aes(x = Power, y = mean.k., label  = Power)) +
  geom_text(color = "red") +
  scale_x_continuous(breaks = powers) +
  scale_y_continuous(transform = "log10", breaks = c(0.01, 0.1, 1, 10, 100, 1000, 10000)) +
  labs(title = paste0("Mean connectivity"), x = "Softpower", y = "Mean connectivity") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))


(p1|p2)/(p3|p4)

ggsave(paste0(outputDir,"/Softpower_plot_DM_",min(powers),"_till_",max(powers),".pdf"), width = 10, height = 10)

# Difference between scale-free topology fit based on normal log2FC or zscaled log2FC
setdiff(soft_data_DM$fitIndices,soft_data_zscaled_DM$fitIndices)

# Select of adjust the softpower used to run the WGCNA model.
softpower = 8 # Modify based on the scale free topology plot results
```


## 2c. DM: Construct the Network and hierarchical clustering
```{r}
# Adjacency matrix using the soft-thresholding power
message("Softpower: ", softpower)

message("Adjacency matrix")
adjacency_matrix_DM = WGCNA::adjacency(datExpr = expression_data_DM,
                                    type = network_type,
                                    power = softpower)


# Topological Overlap Matrix (TOM)
message("TOM")
tom_DM <- WGCNA::TOMsimilarity(adjMat = adjacency_matrix_DM,
                            TOMType = network_type,
                            verbose = verbose)

message("disTOM")
diss_tom_DM <- 1 - tom_DM

# Cluster genes using hierarchical clustering
message("flashClust")
geneTree_DM <- fastcluster::hclust(as.dist(diss_tom_DM), method = method)
dendro_DM <- flashClust::flashClust(as.dist(diss_tom_DM), method = method)

plot(geneTree_DM, main = "Gene Clustering on TOM-based Dissimilarity", sub = "", xlab = "", cex = 0.6)
plot(dendro_DM, main = "Gene Clustering on TOM-based Dissimilarity", sub = "", xlab = "", cex = 0.6)

identical(round(geneTree_DM$height,3),round(dendro_DM$height,3))
```


## 3c. DM: Module Identification using Dynamic Tree Cut
```{r}
# Step 1: Module Identification using Dynamic Tree Cut
# Use dynamic tree cut to identify modules
message("cutreeDynamic")

dynamicMods_DM <- cutreeDynamic(dendro = geneTree_DM, 
                             distM = diss_tom_DM,
                             deepSplit = deepsplit, 
                             pamRespectsDendro = FALSE,
                             minClusterSize = min_module_size,
                             verbose = verbose)

module_table_premerge_DM <- as.data.frame(table(dynamicMods_DM))

# Convert numeric labels into colors for visualization
dynamicColors_DM <- labels2colors(dynamicMods_DM)
table(dynamicColors_DM)

# Plot dendrogram with module colors
plotDendroAndColors(geneTree, dynamicColors_DM, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)


# Step 2: Calculate Module Eigengenes
MEs_DM <- moduleEigengenes(expression_data_DM, colors = dynamicMods_DM)$eigengenes

# Calculate the dissimilarity of module eigengenes
ME_diss_DM <- 1 - cor(MEs_DM)

# Hierarchical clustering of module eigengenes
ME_tree_DM <- hclust(as.dist(ME_diss_DM), method = "average")
plot(ME_tree_DM, main = "Clustering of Module Eigengenes", xlab = "", sub = "")


# Step 3: Merge Similar Modules (module eigengenes with correlation > 0.8)
message("mergeCloseModules")

mergedModules_DM <- mergeCloseModules(exprData = expression_data_DM, 
                                   colors = dynamicMods_DM, 
                                   cutHeight = cutheight,
                                   verbose = verbose,
                                   iterate = FALSE)

# Get the new merged module colors
mergedColors_DM <- mergedModules_DM$colors

module_table_postmerge_DM <- as.data.frame(table(mergedColors_DM))

# Get the new merged module eigengenes
mergedMEs_DM <- mergedModules_DM$newMEs

# Plot dendrogram with merged module colors
plotDendroAndColors(geneTree, cbind(dynamicColors_DM, mergedColors_DM),
                    c("Original Modules", "Merged Modules"),
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)

eg_score_mergedMEs_summary = mergedMEs_DM |> summary() |> as.data.frame() |> select(-Var1) |> 
  separate(col = Freq, into = c("Stat", "Value"), sep = ":") |> 
  pivot_wider(names_from = Stat, values_from = Value)


# Step 4: Get module definition
module_definition_DM <- data.frame(module_premerge = paste0("module_",dynamicMods_DM),
                                module_postmerge = paste0("module_",mergedColors_DM),
                                entrez_id = colnames(expression_data_DM),
                                softpower = softpower,
                                deepsplit = deepsplit,
                                minclustersize = min_module_size,
                                cutheight = cutheight) |>
  dplyr::left_join(y = input |>
                     mutate(entrez_id = paste0("id_", geneid)) |> 
                     dplyr::select(gene_symbol, entrez_id), by = "entrez_id")

module_table_merging_DM <- module_definition_DM |> 
  select(module_premerge, module_postmerge) |>
  mutate(module_nr = as.integer(str_remove_all(module_postmerge, "module_"))) |> 
  table() |> as.data.frame() |> 
  filter(Freq > 0) 
```


# Module Preservation
```{r, include=FALSE}
## Preservation merged modules
modules_compare = input |>
  select(gene_symbol, geneid, probeset) |>
  mutate(entrez_id = paste0("id_", geneid)) |>
  left_join(module_definition |> 
              select(entrez_id, module_premerge, module_postmerge), 
            by = "entrez_id") |>
  left_join(module_definition_TG |> 
              select(entrez_id, module_premerge_TG = module_premerge, module_postmerge_TG = module_postmerge), 
            by = "entrez_id") |>
  left_join(module_definition_DM |> 
              select(entrez_id, module_premerge_DM = module_premerge, module_postmerge_DM = module_postmerge), 
            by = "entrez_id")

data = list(modules_all = as.data.frame(expression_data),
            modules_TG = as.data.frame(expression_data_TG),
            modules_DM = as.data.frame(expression_data_DM))

data = list2multiData(data)

modules = list(modules_all = modules_compare$module_postmerge,
               modules_TG = modules_compare$module_postmerge_TG,
               modules_DM = modules_compare$module_postmerge_DM)

module_preservation = modulePreservation(multiData = data, 
                                         multiColor = modules, 
                                         verbose = verbose,
                                         dataIsExpr = TRUE,
                                         nPermutations = 200,
                                         maxModuleSize = 1000, 
                                         maxGoldModuleSize = 1000,
                                         networkType = network_type,
                                         parallelCalculation = TRUE)

Cross_tabulation_overlap = WGCNA::overlapTable(modules_compare$module_postmerge,modules_compare$module_postmerge_TG)

modules_cross_tabulation_all_vs_TG = Cross_tabulation_overlap$countTable |>
  as.data.frame() 

modules_cross_tabulation_all_vs_TG = modules_cross_tabulation_all_vs_TG |>
  rownames_to_column("module") |>
  pivot_longer(cols = where(is.numeric),names_to = "module_TG",values_to = "overlap") |>
  filter(overlap != 0) |>
  mutate(module_nr = as.integer(str_remove_all(module, "module_"))) |>
  arrange(module_nr, desc(overlap))

Cross_tabulation_overlap = WGCNA::overlapTable(modules_compare$module_postmerge,modules_compare$module_postmerge_DM)

modules_cross_tabulation_all_vs_DM = Cross_tabulation_overlap$countTable |>
  as.data.frame() 

modules_cross_tabulation_all_vs_DM = modules_cross_tabulation_all_vs_DM |>
  rownames_to_column("module") |>
  pivot_longer(cols = where(is.numeric),names_to = "module_TG",values_to = "overlap") |>
  filter(overlap != 0) |>
  mutate(module_nr = as.integer(str_remove_all(module, "module_"))) |>
  arrange(module_nr, desc(overlap))
```


# Module Preservation evaluation
```{r}
module_Zsummary_TG <- module_preservation$preservation$Z$ref.modules_all$inColumnsAlsoPresentIn.modules_TG |> rownames_to_column("module")

module_tabulation_TG <- module_preservation$accuracy$observedCounts$ref.modules_all$inColumnsAlsoPresentIn.modules_TG |>
  t() |> as.data.frame() |> 
  rownames_to_column("module")

module_Zsummary_DM <- module_preservation$preservation$Z$ref.modules_all$inColumnsAlsoPresentIn.modules_DM |> rownames_to_column("module")

module_tabulation_DM <- module_preservation$accuracy$observedCounts$ref.modules_all$inColumnsAlsoPresentIn.modules_DM |>
  t() |> as.data.frame() |> 
  rownames_to_column("module")

module_Zsummary <- module_Zsummary_TG |>
  left_join(module_Zsummary_DM, by = c("module", "moduleSize")) |>
  mutate(moduleSize_log10 = log10(moduleSize))

colnames(module_Zsummary) <- gsub("[.]x", ".TG",colnames(module_Zsummary))
colnames(module_Zsummary) <- gsub("[.]y", ".DM",colnames(module_Zsummary))

ggplot(module_Zsummary, aes(x = Zsummary.pres.TG, y = Zsummary.pres.DM, colour = moduleSize_log10, group = module)) +
  geom_point() + 
  theme_bw() +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1) +
  geom_hline(yintercept = c(2,10), colour = c("orange", "green")) +
  geom_vline(xintercept = c(2,10), colour = c("orange", "green")) +
  scale_color_gradientn(colours = rev(brewer.pal(9,"Spectral")))

ggplotly(width = 1000,height = 1000)
```



# Save data
```{r, include=FALSE}
save(input, expression_data, expression_data_DM, expression_data_TG, module_definition, module_definition_DM, module_definition_TG,
     file = file.path(outputDir, "WGCNA_TG_DM_rat_kidney_preservation_input.RData"))

save(data, modules, modules_compare, module_preservation, modules_cross_tabulation_all_vs_DM, modules_cross_tabulation_all_vs_TG,
     module_Zsummary_TG, module_Zsummary_DM, module_tabulation_TG, module_tabulation_DM,
     file = file.path(outputDir, "WGCNA_TG_DM_rat_kidney_preservation_output.RData"))

save.image(file.path(outputDir, "WGCNA_TG_DM_rat_kidney.RData"))
```

